\chapter{Universal Algebra: An Overview}
By the early 18th century, mathematicians had discovered how to solve polynomial
equation of up-to degree 4. While trying to find a general solution for
polynomial equation, Lagrange and Abel established permutation groups
\cite{barnett2017roots}. These developments including in the field of algebraic
structures and geometry became the main source for \textit{group theory}
\cite{enwiki:1145792803}. It was mathematician Evariste Galois who coined the
term \emph{group} and established group theory. He used group to determine the
solvability of polynomial equations. Group theory was later discovered to be
useful in various fields of mathematics such as modulus theory and geometry
\cite{enwiki:1107380309}. As group theory, the study of group structures
evolved, other abstract structures were invented to solve different problems.
This gave rise to a new field in mathematics called \emph{abstract algebra}.
Abstract algebra is the study of algebraic structure and its models or examples
\cite{enwiki:1107380309}. An algebraic structure is a tuple containing a
carrier set $A$, a set of operations that act on $A$, and a set of axioms
involving the operations on $A$.

Algebraic structures, like monoids, loops, groups, and rings have similar
properties. Universal algebra studies these structures by abstracting out the
specific definitions and properties of algebraic structures. Universal algebra
will deal with these algebraic structures as axiomatic theories in equational
first-order logic \cite{YSharoda}. The definitions in this chapter are adapted
from \cite{sankappanavar1981course}, \cite{YSharoda}, and
\cite{sannella2012foundations}.

\section{Relation and function}
In order to understand algebraic structures, it is essential to know some basics
of relations and functions. In this section, we define relations and functions.
We can start with defining a set.
\begin{itemize}
\item A set is a well-defined collection of objects. The elements or members of the
set can be mathematical objects of any kind such as numbers, symbols, geometrical
shapes, or even other sets. If $x$ is an object in set $S$ then we say x is an
element of S and is denoted as $x \in S$.

\item The \emph{Cartesian product} between two sets $X$ and $Y$,  $X \ \times\
Y$ is defined as the set of pairs \(\{(x,y) :\ x \ \in\ X,\ y\ \in\ Y \}\). 

\item A \emph{binary relation}  $R$ over sets $X$ and $Y$ is a subset of the
Cartesian product $X \times Y$. Set $X$ is called the domain and set $Y$ is
called the codomain of $R$.

\item A relation $R$ is equivalence if it satisfies:
\begin{enumerate}
    \item Reflexive: A \emph{reflexive relation} $R$ on set $X$ is a subset on
\(X \times X\) and $\forall x \in S, (x,x) \in R$

    \item Symmetric: A \emph{symmetric relation} R on set X is a subset of \(X
\times X\) such that \(\forall x y \in X,\ (x,y) \in R ⟺ (y,x) \in R \)

    \item Transitive: A \emph{transitive relation} on set X is a subset of \(X
\times X\) such that \(∀ x y z \in X \), if \((x,y) \in R\) and \((y,z) \in R \)
then \((x,z) \in R\)
\end{enumerate}
In other words, a relation R is an \emph{equivalence} if it is reflexive, symmetric and transitive.

\item In a relation, if every element in the domain $(X)$ is mapped to only one
element in the codomain $(Y)$, then we call it a \emph{function}. This can be
expressed using the notation:
\[f:\ X \rightarrow\ Y\]
\[x \ \mapsto\ f(x)\]
For example, on the set of natural numbers $N$ to $N$ we can define a function as:
\[f:\ N \rightarrow N\]
\[x \ \mapsto\ x^{2}\]   
\item Let $ X $ and $ Y $ be two sets, and $ f:X \rightarrow Y$ be a function then:
\begin{enumerate}
    \item The function $f$ is \textit{identity} if $X=Y$ and $f(x)\ =\ x,\
    \forall x\ \in \ X$.
    \item A function $f$ is \emph{injective} if $f$ maps distinct elements of
    the domain to distinct elements of the codomain. $f$ is injective if
    $f(x)=f(y) \Rightarrow x = y, \forall x,y \in X.$
    \item A function is called \emph{surjective} if given $y \in Y$, there
    exists $x\in X$ such that $f(x) = y$.
    \item A function is called \emph{bijective} if it is both injective and
    surjective.
\end{enumerate}
\item An \emph{operation} is defined as a function that can take zero or more inputs
and maps them to a well-defined output value. The number of operands is the arity
of the operation.
\end{itemize}

\section{Universe, type, and signature}
Logic allows to describe properties of entities as formulas and provide
reasoning about them. Equational logic limits these formulas (such as axioms or
theorems) to  be universally quantified equations of the form $t_1 = t_2$. Here
$t_1$ and $t_2$ are terms expressible in the language of theory. A proposition
is true if it is derivable from other true propositions using inference rules.
The three inference rules in equational logic described in
\cite{gries2013logical} are:
\begin{itemize}
    \item Leibniz equality: Two expressions are equal if one expression can be
    substituted with other without changing the truth statement. \[\frac{t_1
    = t_2}{t[x \mapsto t_1] = t[x \mapsto t_2]}\]
    \item Transitivity: If $t_1 = t_2$ and $t_2 = t_3$ then $t_1 = t_3$.
    \[{\frac{t_1 = t_2\ t_2 = t_3}{t_1 = t_3}}\]
    \item Substitution: For predicate $p$, if $p\ t$ is true, it remains true on all
    conditions. \[\frac{p t}{p(t[xs \mapsto ts])}\]
\end{itemize}
where $t,t_1,t_2, \text{and} t_3$ are term expressions, $x$ is some symbol in the
language, $xs$ and $ts$ denotes list of symbols and list of expressions
respectively. 

A theory in universal algebra is defined as a tuple $(S,F,E)$ such that, $S$ is
a set of single sort $s$, $F$ is a finite set of function symbols with their
arities, and $E$ is a set of equations that are satisfied in $S$.

A sub theory of a theory $(S,F,E)$ is defined as $(S_\Delta,F_\Delta,E_\Delta)$
such that $S_\Delta \subseteq S$. The operations in sub theory $(op_\Delta \in
|F_\Delta|)$ is defined as
\[op_\Delta\ x_1...x_n = op\ x_1...x_n \in S_\Delta,\forall op \in
|F|,\text{and}\ x_1...x_n \in S_\Delta \]  

A signature of a theory $(S,F,E)$ is a pair ($S,F$) such that $S$ is set of
sorts and $F$ is set of operation names. An operation can be defined as a triple
$(f,[s_1,...,s_n],s)$ consisting of name, its arity, and the target sort. 

An algebra $A$ is defined as $A = (S_A,F_A)$ a mathematical structure consisting
of a domain and functions on the domain. Algebra provides an interpretation for
the carrier set $S$ and function symbols $F$ of a theory.

For example, a group G is an algebra with one nullary (1), one unary ($^{-1}$),
and one binary (∙) operation represented as $(G, ∙, ^{-1}, 1)$. Here, $^{-1}$ is
an inverse operation. For a binary operation ∙, on set $S$, if $x ∙ y = e$ where
$e$ is an identity element, and $x,y \in S$ then $x$ is called the left inverse
of $y$, and $y$ is called the right inverse of $x$. An element is said to have
an inverse if it has both left and right inverse. The group structure should
satisfy the following axioms. 
\begin{enumerate}
\item Associativity - \( ∀ x y z \in G, x ∙ (y ∙ z) = (x ∙ y) ∙ z \)
\item Identity - \(∀ x \in G, x ∙ 1 = 1 ∙ x = x\)
\item Inverse - \( ∀ x \in G, x ∙ x ^{-1} =  x
^{-1} ∙ x = 1\)
\end{enumerate}

The type (or language) of the algebra is a set of function symbols. Each member
of this set is assigned a positive number which is the arity of the member. For
example, an algebra of type (2,0) denotes an algebra with one binary operation
and one nullary operation. The group structure defined above is of type (2,1,0).
That is $∙$ is a binary operation, $^{-1}$ is a unary operation and $1$ is the
nullary operation.

\section{Constructions}
Universal algebra provides definitions of constructions related to algebraic
structures. In this section, we will describe some of these constructions. 
\begin{itemize}
    \item The \textit{congruence} relation for an algebraic structure can be
    defined as an equivalence relation that is compatible with the structure
    such that the operations are well-defined on the equivalence class. For an
    algebra $(A,F)$, $\theta$ is a congruence on $A$ if $\theta$ satisfies the
    compatibility property. The compatibility property states that for each
    n-ary function symbol $f \ \in\ F$ and $x_i,\ y_i\ \in\ A$, If $x_i\ \theta\
    y_i$ holds for \(1\leq i \leq n\) then $f^{A}(x_1,...,x_n)\ \theta\
    f^{A}(y_1,....,y_n)$ holds \cite{sankappanavar1981course}.

    For example, consider group structure $(G, ∙, ^{-1}, 1)$. A
    congruence relation on $G$ with binary operation $∙$ is an equivalence
    relation $\equiv$ on $G$ such that \[g_1\equiv g_2\ \text{and}\ h_1 \equiv h_2
    \Rightarrow g_1 ∙ h_1 \equiv g_2 ∙ h_2\ \forall g_1,g_2,h_1,h_2 \in G\] 

    \item A \textit{morphism} is a structure preserving map between two
    algebraic structures. It is an abstraction that generalizes the map between
    two structures or mathematical objects in general. If $A$ and $B$ are two
    algebras of same type $F$, then a homomorphism is defined as a function
    $\alpha: A \rightarrow B$ such that: \[ \alpha\ (f^{A}(a_1....a_n))\ =\
    f^{B}\ ((\alpha\ a_1)....(\alpha\ a_n))\] For each n-ary $f$ in $F$ and
    each sequence $a_1....a_n$ from $A$.

    Some variants of homomorphism are:
    \begin{enumerate}
        \item  Monomorphism: For two algebras $A$ and $B$, if \(\alpha : A
        \rightarrow B \) is a homomorphism from $A$ to $B$, and if \(\alpha\)
        satisfies one-to-one mapping (i.e., \(\alpha\) is injective) then the
        morphism \(\alpha\) is called a \textit{monomorphism}.

        \item Isomorphism: For algebra $A$ and $B$, a homomorphism $f:A → B$ is
        an isomorphism if it has an inverse, i.e. there is a homomorphism
        $f^{-1}:B → A$ such that $ff^{-1} =
        id_{|A|}$ and $f^{-1}f = id_{|B|}$ 

        \item Endomorphism: A homomorphism from an algebra $A$ to itself is
        called \textit{endomorphism}. In other words, if $f$ is a homomorphism on $A$
        such that $f:A\rightarrow A$ then, f is an endomorphism.

        \item Automorphism: An isomorphism from an algebra $A$ to itself is
        called \textit{automorphism}.

        \item Epimorphism: For two algebras $A$ and $B$, if \(\alpha : A
        \rightarrow B \) is a homomorphism from $A$ to $B$, and if \(\alpha\) is
        surjective then the morphism \(\alpha\) is called a
        \textit{epimorphism}.
    \end{enumerate}

    \item For algebras $A$, $B$, and $C$ the \textit{composition of morphisms}
    $f:\ A \ \rightarrow \ B$ and $g:\ B \rightarrow\ C$ is denoted by the
    function $g\ \circ \ f\ :\ A\ \rightarrow \ C$ and is defined as $(g\ \circ
    \ f)\ a = \ g\ (f\ a), \ \forall \ a\ \in\ A$. In
    \cite{sankappanavar1981course}, the author proves that the composite of two
    homomorphisms (monomorphisms/isomorphisms) is also a homomorphism
    (monomorphism/isomorphism).

    \item A quotient algebra for some theory $(S,F,E)$ with respect to
    congruence relation $(\cong)$ is defined on the theory $(S_Q,F_Q,E_Q)$ where
    $S_Q$ is the factor set of $S$ such that: \[S_Q = \{[x] | x \in S\}\] The
    operations $op_Q$ in the quotient algebra is defined as: \[op_Q\
    [x_1]...[x_n] = [op\ x_1 ... x_n]\] where $op_Q \in F_Q$ and $op \in F$, and
    $[x]$ denotes the equivalence class such that \([x] = \{y \in S | x \cong
    y\}\).
    
    \item \textit{Direct product}: For set of algebra $\{ A_i | i \in I \}$ of
    same type indexed by some arbitrary set $I$, the cartesian product of the
    underlying sets is defined as $A = \displaystyle \prod_{i\in I} A_i$. Let
    $\omega_{A_i}$ be the corresponding n-ary operator on $A_i$. We can define
    $\omega_A : A^{n} \rightarrow A$ by \[ \omega_A(a_1,...a_n)(i) =
    \omega_{A_i} (a_1(i),...,a_n(i)) \forall i \in I\] where element $a \in A$
    is a function from indexing set $I$ to $\bigcup A_i$ such that $i \in I,
    a(i) \in A$. The algebra $A$ equipped with all $\omega_A$ on $A$ is the
    direct product of $A_i$. Each $A_i$ is called the direct factor of $A$.

\end{itemize}